Gastrointestinal II

### 2017-09-28 git2.rmd
sc

Coeliacs: gluten => gliadin (in the stomach; resists breakdown) => reaches the SI then binds to TFR (transferrin receptor; usually used to absorb iron, may be overexpressed) => transcytosed across enterocyte => lamina propia => gliadin is deamidated then phagocytosed by macrophages => macros are usually doing this for normal food anyway; abnormal reaction in coeliacs => Th2 (CD4+) => inflamm cytokines such as interferon-$\gamma$ and TNF-$\alpha$ => direct damage to epithelial cells in SI + killer CD8 t-cells stimulated => bad. 
Ab blood tests: anti-tTG (deamidates the gliadin) and anti-endomysial (EMAs). 
Symptoms = {children = [abd distension, failure to thrive, diarrhoea], adults = [chronic diarrhoea, bloating (distension),...]}. 
Tl;dr autoimmune.

@@

-  Microcytic = {T = thalassaemia, I = iron deficiency, C = chronic disease, S = sideroblastic anaemia}
-  Macrocytic = {A = alcohol (cirrhosis), B = B12 deficiency, C = compensatory reticulocytosis (blood loss and hemolysis), D = drug/dysplasia, E = endocrine (hypothyroidism), F = folate deficiency/foetus (pregnancy)}

@@

@@

>>>>

Epidemiology: caucasians of european ancestry.

Coeliacs: gluten => gliadin (in the stomach; resists breakdown from gastric enzymes) => reaches the SI then binds to TFR (transferrin receptor; usually used to absorb iron, may be overexpressed) => transcytosed across enterocyte => lamina propia => gliadin is deamidated then phagocytosed by macrophages => macros are usually doing this for normal food anyway; abnormal reaction in coeliacs => Th2 (CD4+) => inflamm cytokines such as interferon-$\gamma$ and TNF-$\alpha$ => direct damage to epithelial cells in SI + killer CD8 t-cells stimulated => bad. 
Ab blood tests: anti-tTG (deamidates the gliadin) and anti-endomysial (EMAs). 
Symptoms = {children = [abd distension, failure to thrive, diarrhoea], adults = [chronic diarrhoea, bloating (distension),...]}. 
Tl;dr autoimmune.

@@

-  Microcytic = {T = thalassaemia, I = iron deficiency, C = chronic disease, S = sideroblastic anaemia}
-  Macrocytic = {A = alcohol (cirrhosis), B = B12 deficiency, C = compensatory reticulocytosis (blood loss and hemolysis), D = drug/dysplasia, E = endocrine (hypothyroidism), F = folate deficiency/foetus (pregnancy)}
-  RACH, BLOOD?

@@

-  Repro/gynae: pregnancy, ovarian cysts/cancer, fibromatosis
-  GIT: hernia, malabsorption (lactose intolerance => osmotic diarrhoea), coeliacs, IBD, infective, peptic ulcer (sequelae include cancers)
-  Psych: depression, chronic stress
-  Systemic: anaemia (Fe deficiency can lead to diarrhoea)
-  Endocrine: hypothyroidism, diabetic ketoacidosis

Ex:

-  Inspection: scarring, swelling, asymmetry, skin changes, guarding, general discomfort, also inspect eyes for jaundice/anaemia
-  Palpation: masses (firm/fixed/flatulent), organs, belot (wiggle) the kideys, for pain
-  Auscultate
-  Percussion: if considering ascites
-  Rectal

Aphthous ulcers: benign mouth ulcers associated with systemic stuff like coeliacs

-  Bloods: FBC (Hb = anaemia), thyroid fn test, HLA typing, B12, ELFT
-  Stool sample
-  Tissue typing: ?
-  Discuss possibility for endoscopy: skip lesion things like Crohns; wait for bloods  maybe

Ix: suggest of coeliacs disease (systemic autoimmune disease). IgA-tTG: deamidating gluten products

Further Ix:

-  Iron studies
-  Gastroscopy and biopsy
-  Diet

Coeliacs associated with FHx of thyroid issues and T1D, Trisomy-21, IBD, GORD

Complications: bloating, chronic diarrhoea, nutritional deficiency, increased risk of malignancy, higher motility

-  Gluten free diet
-  Frequent monitoring of coeliacs and systemic nutrient levels

Gluten found in wheat, rye, HLA-DQ2 (95%) and HLA-DQ-8. Can present at any age and common (1%). Learn the histopath => GIT systems + diverse systemic symptoms

@@

-  Hernias: read medscape article
-  Cancer

-  Reducible/irreducible, strangulated/incarcerated (soft/hard)
-  Fluctuants: cyst can look like hernia
-  Worse

Lump:

-  Fixed is worse than mobile

Mx: surgery

Natural Hx: will only get worse

The hernia became strangulated => ischamia => bowel necrosis

Medical emergency

Vomiting:

-  Pain
-  Bowel obstruction
